---
title: "Lab 5 Team Document"
output: html_document
author: Mario Lloyd Galvao-Wilson, Justin Park, and Brian Sayre
---


### Data Input and Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
library(readxl)

acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

prsn <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```

### Question 1
```{r}
acc$Weekday <- wday(acc$DAY_WEEK, label = T)

acc %>%
  ggplot(aes(Weekday)) + geom_bar() + ylab('Number of Accidents') + ggtitle('Accidents per Day of the Week')
```

The most accidents occur on Fridays, Saturdays, and Sundays. Saturdays have the most accidents of any day of the week, which the number of accidents on Fridays and Sundays are comparable. 

### Question 2
```{r}
fatal <- prsn %>%
  filter(INJ_SEV == 4)
```

### Question 3
```{r}
q3 <- fatal %>%
  group_by(STATE, MAKE) %>%
  summarise(nf = n())

danger <- q3 %>% 
  na.omit() %>%
  group_by(STATE) %>%
  filter(nf == max(nf))
```

### Question 4
```{r warning=FALSE}
codes <- readxl::read_xlsx("data/state_codes.xlsx")
states <- map_data('state')

codes$'State Name' <- tolower(codes$'State Name')
codes$'State Code' <- as.numeric(codes$'State Code')

labels <- states %>%
   group_by(region) %>%
   summarise(long= mean(long, na.rm = TRUE), lat= mean(lat, na.rm = TRUE))

labels <- left_join(labels, codes, by = c("region" = "State Name"))

labels <- left_join(labels, danger, by = c("State Code" = "STATE"))

labels$MAKE <- factor(labels$MAKE, levels = c(12, 20, 49, 37), 
                     labels = c("Ford", "Chevrolet", "Toyota", "Honda"))                         
states %>% ggplot() +
  geom_polygon(  aes(x=long, y=lat, group=group),
                color="grey", fill="black" ) + geom_text(data=labels, aes(label = MAKE, x=long, y=lat), color="white", size=3)

```

From the map that we made I see a trend where the most dangerous cars in the East tend to be Chevy and as you go more West it starts to shift to Ford. Overall I think there is more Chevy than Ford on this map.

### Question 5
```{r}
q5 <- inner_join(prsn, acc, by="ST_CASE")
```

### Question 6
```{r}
q6 <- q5 %>%
  filter(SEX == 1 || SEX == 2) %>%
  group_by(ST_CASE) %>%
  summarise(avgSex = mean(SEX))

q6$sexStr <- ifelse(q6$avgSex == 1, 'Men', ifelse(q6$avgSex == 2, 'Women', 'Both'))

accident <- q6 %>%
  select(ST_CASE, sexStr) %>%
  inner_join( acc, by='ST_CASE')

grouped <- accident %>%
  filter(HOUR <= 24) %>%
  filter(DAY_WEEK != 9) %>%
  group_by(sexStr, HOUR, DAY_WEEK) %>%
  summarise(numAccidents = n()) %>%
  arrange(desc(numAccidents))

ggplot(grouped,aes(x=HOUR, y=numAccidents)) + 
  geom_bar(stat='identity') + 
  facet_grid(sexStr~DAY_WEEK) + 
  xlab("Hours of the day") + 
  ylab("Total number of accidents")
```

In terms of sex I found that both men and women accidents are the most common and men are right behind. You can see that women is much lower than both other categories. In terms of the hour of the day we can see that accidents usually peak later at night and really dip in the mornings.

```{R}
glc <- read_xlsx("FRPP_GLC_-_United_StATESaPRIL62021.xlsx")
states <- map_data('state')
counties <- map_data('county')
```

### Question 7
```{R}
myvars <- c("STATE", "COUNTY", "FATALS")
fatalities <- acc[myvars]

glc$'State Code' <- as.numeric(glc$`State Code`)
glc$'City Code' <- as.numeric(glc$`City Code`)
glc$'County Code' <- as.numeric(glc$`County Code`)

fatalities <- fatalities %>%
  group_by(STATE) %>%
  group_by(COUNTY) %>%
  summarize(FATALS = sum(FATALS))

fatalities <- fatalities %>% left_join(glc, by=c('COUNTY' = 'County Code'))

fatalities$`State Name` <- tolower(fatalities$`State Name`)
fatalities$`County Name` <- tolower(fatalities$`County Name`)

fatalities <- fatalities %>% right_join(counties, by=c('County Name' = 'subregion', 'State Name' = 'region'))

ggplot(fatalities, aes(x=long, y=lat)) + 
  geom_polygon(aes(group=group)) +
  geom_polygon(aes(group=group, fill=FATALS))
```

There are some large areas on the map that have "holes" where there have been no reported fatal car accidents. This makes sense since certain parts of the country do not have a lot of traffic such as mountains in Colorado, New Mexico, and Arizona. You would also not expect to see many fatal car accidents in cities since people do not drive very fast in areas there is a lot of traffic.

### Question 8
```{R}
Winter <- acc %>%
          filter(MONTH  < 4 | MONTH > 9)

Summer <- acc %>%
          filter(MONTH > 4 & MONTH < 9)

winterFatal <- Winter %>%
  group_by(STATE) %>%
  summarize(FATALS = sum(FATALS))

summerFatal <- Summer %>%
  group_by(STATE) %>%
  summarize(FATALS = sum(FATALS))

summerFatal <- summerFatal %>% left_join(glc, by=c('STATE' = 'State Code'))
summerFatal$`State Name` <- tolower(summerFatal$`State Name`)
summerFatal <- summerFatal %>% left_join(states, by=c('State Name' = 'region'))

myvarstwo <- c("State Name", "FATALS")
summerFatal <- summerFatal[myvarstwo]

summerFatal <- unique(summerFatal)

ggplot(data = summerFatal, aes(x=`State Name`, y=FATALS)) + geom_bar(stat = 'identity') + labs(title = "Accidents by State in the Summer", y = "Number of Fatalities")+ 
  theme(axis.text.x = element_text(angle = -90, vjust = 0, hjus=0))

winterFatal <- winterFatal %>% left_join(glc, by=c('STATE' = 'State Code'))
winterFatal$`State Name` <- tolower(winterFatal$`State Name`)
winterFatal <- winterFatal %>% left_join(states, by=c('State Name' = 'region'))

myvarstwo <- c("State Name", "FATALS")
winterFatal <- winterFatal[myvarstwo]

winterFatal <- unique(winterFatal)

ggplot(data = winterFatal, aes(x=`State Name`, y=FATALS)) + geom_bar(stat = 'identity') + labs(title = "Accidents by State in the Winter", y = "Number of Fatalities") + 
  theme(axis.text.x = element_text(angle = -90, vjust = 0, hjus=0)) 
```
I declared winter as October through March and summer as May through August. Winter seems to be more dangerous since there are more car accidents in each state in the Winter. In general whether it is summer or winter, the same states tend to have the most and the least number of fatal car accidents. Texas and California have the most fatal car accidents which makes sens since they are the largest states and there is more traffic.